---
title: "Tracking Absentee Ballot Status"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
knit: (function(i, encoding) {apstyle::render(i, encoding)})
---
<!-- Timestamp: leave this in to automatically track when a file is rendered -->
*Rendered at `r format(Sys.time(), '%I:%M %p, %b. %d, %Y')`*

<!-- Contact info: leave author/reporter/editor contact info here -->
*Ryan Thornburg (<ryan.thornburg@unc.edu>)*

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
  )

knitr::opts_knit$set(
  root.dir = rprojroot::find_rstudio_root_file()
  )
```

```{r initialize}
# Load files, environment variables, libraries, etc. here
library(tidyverse)
library(lubridate)
library(janitor)
library(apstyle)
library(DT)
```


## Loading Data
```{r loading, message = FALSE}

temp <- tempfile()
filename_daily<- paste0("absentee_20201103","_", format(Sys.time(), '%y%m%d'),".csv")
download.file("https://s3.amazonaws.com/dl.ncsbe.gov/ENRS/2020_11_03/absentee_20201103.zip", temp <- temp)
unzip(temp, exdir = "data")

file.rename(file.path("data","absentee_20201103.csv"), file.path("data",filename_daily))


absentee_20201103 <- read_csv(file.path("data",filename_daily), 
                              col_types = cols(voter_zip = col_character(), 
                                               other_city_state_zip = col_skip(), 
                                               relative_request_name = col_skip(), 
                                               relative_request_address = col_skip(), 
                                               relative_request_city = col_skip(), 
                                               relative_request_state = col_skip(), 
                                               relative_request_zip = col_skip(), 
                                               election_dt = col_date(format = "%m/%d/%Y"), 
                                               ballot_req_dt = col_date(format = "%m/%d/%Y"), 
                                               ballot_send_dt = col_date(format = "%m/%d/%Y"),
                                               ballot_rtn_dt = col_date(format = "%m/%d/%Y")))

```

## Create Status Tracker, by NCID
```{r pivot}

ncid_status_tracker<- absentee_20201103 %>%
  #On Oct. 15 20 unique voters had multiple rows for the same ballot_rtn_status + ballot_rtn_date. At least one had multiple dates for the same ballot return status.
  group_by(ncid, voter_reg_num, ballot_rtn_status, ballot_rtn_dt) %>%
  #Use values_fn = max here to take the most recent date if a voter has two dates for the same status.
  pivot_wider(names_from = ballot_rtn_status, values_from = ballot_rtn_dt, values_fn = max) %>%
  clean_names() %>%
  #This line only works if there are exactly 13 valid values in the original ballot_rtn_status field
  select(ncid, voter_reg_num,accepted, accepted_cured, pending_cure, spoiled, returned_undeliverable, witness_info_incomplete, pending, cancelled, conflict, duplicate, signature_different, other=na, e_transmission_failure) 

tracker_filename <- paste0("ncid_status_tracker", format(Sys.time(), '%y%m%d'), ".csv")


rio::export(ncid_status_tracker, file.path("data",tracker_filename))

#datatable(ncid_status_tracker)
```

## Days Between Witness Error and Acceptance
```{r duration}

duration <- ncid_status_tracker %>%
  filter(!is.infinite(witness_info_incomplete)) %>%
  filter(!is.infinite(accepted) | !is.infinite(accepted_cured)) %>%
  #I'm not sure about the different meanings of "accepted" vs. "accepted cured" but they are mutually exclusive.
  mutate(accepted = case_when(is.infinite(accepted) ~ accepted_cured,
                                !is.infinite(accepted) ~ accepted))


duration_filename <- paste0("witness_accepted", format(Sys.time(), '%y%m%d'), ".csv")

rio::export(duration, file.path("data",duration_filename))

#datatable(duration)
```
As of `r format(Sys.time(), '%B %d')` there were `r absentee_20201103 %>% filter(ballot_rtn_status=="WITNESS INFO INCOMPLETE") %>% nrow()` vote-by-mail ballots that had county board of election said had incomplete witness information. But `r duration %>% nrow()` were later accepted. 

## Average Time from Witness Incomplete to Acceptance
```{r avg_duration}

avg_duration<- duration %>% as_tibble() %>% mutate(lag = accepted - witness_info_incomplete) %>% select(witness_info_incomplete, accepted, lag) %>% summarize(avg_lag= mean(lag))

avg_duration<-round(avg_duration$avg_lag)
```
The average time between the time a county board finds a ballot with an incomplete witness is `r avg_duration` days.



## References

<!-- This section is for including references to related material -->

- https://twitter.com/OldNorthStPol/status/1315346895779368968/photo/1

- https://www.ncsbe.gov/results-data/absentee-data
